\LARGE \bf Steering a Swarm Using Global Inputs and Swarm Statistics

### 2016-08-12 journalSwarmControl/relatedWork.tex
finished Becker edits to entire paper

\subsection{Block pushing and compliant manipulation}
Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object, as in \cite{Sudsang2002,Fink2007}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object. A key difference is that our robots are compliant and tend to flow around the object, making this similar to fluidic trapping as in \cite{Armani2006} and \cite{Becker2009}.

>>>>

\subsection{Block pushing and compliant manipulation}
Unlike \emph{caging} manipulation, where robots form a rigid arrangement around an object, as in \cite{Sudsang2002,Fink2008}, our swarm of robots is unable to grasp the blocks they push, and so our manipulation strategies are similar to \emph{nonprehensile manipulation} techniques, e.g.~\cite{Lynch1999}, where forces must be applied along the center of mass of the moveable object. A key difference is that our robots are compliant and tend to flow around the object, making this similar to fluidic trapping as in \cite{Armani2006} and \cite{Becker2009}.

### 2016-08-12 journalSwarmControl/simulation.tex
Changed Sections and u
Latest changes
finished Becker edits to entire paper

We performed a parameter sweep using the PD controller \eqref{eq:PDcontrolPosition} to identify the best control gains.  Representative experiments are shown in Fig.~\ref{fig:gainvalues}. 100 robots were used and the maximum speed was 3 meters per second. As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52 s with $K_{p}= 4$, and  $K_{d} = 1$. 
\begin{figure}
\centering
\begin{overpic}[width = 0.7\columnwidth ]{gains.eps}
\end{overpic}
%\begin{overpic}[width =0.49\columnwidth]{meanVariance4.eps}
%\end{overpic}
\vspace{-1em}
\caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots. %b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom). For ease of analysis, only $x$ position and variance are shown.
%\vspace{-2em}
}
\end{figure}

@@

For variance control we use the control law discussed in Section~\ref{sec:VarianceControl}. 
 Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.

>>>>

We performed a parameter sweep using the PD controller \eqref{eq:PDcontrolPosition} to identify the best control gains.  Representative experiments are shown in Fig.~\ref{fig:gainvalues}. 100 robots were used and the maximum speed was 3 meters per second. As shown in Fig.~\ref{fig:gainvalues}, we can achieve an overshoot of 1\% and a  rise time of 1.52~s with $K_{p}= 4$, and  $K_{d} = 1$. 
\begin{figure}
\centering
\begin{overpic}[width = 0.7\columnwidth ]{gains.eps}
\end{overpic}
%\begin{overpic}[width =0.49\columnwidth]{meanVariance4.eps}
%\end{overpic}
\vspace{-1em}
\caption{\label{fig:gainvalues} In simulation, tuning proportional ($K_p$, top) and derivative ($K_d$, bottom)  gain values in \eqref{eq:PDcontrolPosition} improves performance with $n = 100$ robots. %b) Simulation result with 100 robots under hybrid control Alg.~\ref{alg:MeanVarianceControl}, which  controls both the mean position (top) and variance (bottom). For ease of analysis, only $x$ position and variance are shown.
%\vspace{-2em}
}
\end{figure}

@@

For variance control we use the control law \eqref{eq:PDcontrolVariance}. 
 Results are shown in Fig.~\ref{fig:varyBrownian}, with $K_{p,i,d} = [4,1,1]$.

### 2016-08-12 journalSwarmControl/theory.tex
finished Becker edits to entire paper

and our state space representation as:
\begin{equation}
\begin{bmatrix}
\dot{x}\\ 
\ddot{x}\\
\dot{y}\\
\ddot{y}
\end{bmatrix} = \begin{bmatrix}
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0\\
0 & 0 & 0 & 1\\
0 & 0 & 0 & 0
\end{bmatrix}  \begin{bmatrix}
x\\
\dot{x}\\
y\\
\dot{y}
\end{bmatrix} + \begin{bmatrix}
0 & 0 \\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{bmatrix}  [u_x,u_y]
\end{equation}

@@

\begin{equation}
\textrm{Here }
\mathcal{C}=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],
\end{equation}
and thus all four states are controllable. This section starts by proving independent position control of many robots is not possible, but the mean position can be controlled. We then provide conditions under which the variance of many robots is also controllable.

\subsection{Independent control of many robots is impossible}
A single robot is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
%\begin{equation}\left[ x_{1,1},x_{2,1},\ldots, x_{2n-1},x_{2n}\right]^\intercal = \left[ p_{x,1},v_{x,1},\ldots,p_{x,n},v_{x,n}\right]^\intercal. \nonumber\end{equation}
%\begin{align}
%\dot{p}_{x1} &= v_{x1}\\\nonumber
%\dot{v}_{x1} &= a_{x1}\\\nonumber
%\dot{p}_{x2} &= v_{x2}\\\nonumber
%\dot{v}_{x2} &= a_{x2}\\\nonumber
%&\vdots\\\nonumber
%\dot{p}_{xn} &= v_{xn}\\\nonumber
%\dot{v}_{xn} &= a_{xn}\nonumber
%\end{align}
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because $\mathcal{C}$ has rank two:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,, 
\ldots \right]
\end{equation}  
\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} 
+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

We again analyze $\mathcal{C}$:
\begin{equation}
\mathcal{C}=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right]
\end{equation}
This matrix has rank two, and thus the average position and average velocity are controllable.

@@

We control mean position with a PD controller that uses the mean position and mean velocity. Our control input is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\bar{v}_x) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\bar{v}_y)  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. 
Microscale systems are affected by unmodelled dynamics. 
These unmodelled dynamics are dominated by Brownian noise, described by~\cite{einstein1956investigations}. 
To model this~\eqref{eq:stdform} must be modified as follows:
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t)
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$. Given a large obstacle-free workspace, a \emph{Brownian noise} process increases the variance linearly with time.
\begin{equation}\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon} \end{equation}
 If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board~\cite{Becker2013b}.

%After some time, Gaussian distribution shapes the outline of the robots because of the Brownian noise feature:\\
%\begin{equation}
%P(x) = \frac{1}{\sigma \sqrt{2\pi } }e^{ -  \frac{\left( x - \mu  \right)^2 } {2\sigma ^2 }}
%\end{equation}
If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

@@

A  controller to regulate the variance to $\sigma^2_{ref}$ is:
\begin{align}
u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}) \nonumber\\
u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
\end{align}
In a slight abuse of notation we call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{goal}(\sigma^2_{ref}) = r + \sqrt{3\sigma^2_{ref}}
\end{align}
 If another wall is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.

@@

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$ 
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$

>>>>

and our state space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}\\ 
\ddot{x}\\
\dot{y}\\
\ddot{y}
\end{bmatrix} = \begin{bmatrix}
0 & 1 & 0 & 0 \\
0 & 0 & 0 & 0\\
0 & 0 & 0 & 1\\
0 & 0 & 0 & 0
\end{bmatrix}  \begin{bmatrix}
x\\
\dot{x}\\
y\\
\dot{y}
\end{bmatrix} + \begin{bmatrix}
0 & 0 \\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{bmatrix} 
 \begin{bmatrix}
 u_x\\
 u_y\end{bmatrix} 
\end{equation}

@@

\begin{equation}
\textrm{Here }
\mathcal{C}=\left[
\begin{matrix} 
0 & 0\\
\frac{1}{m} & 0 \\
0 & 0 \\
0 & \frac{1}{m}
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
\frac{1}{m}& 0\\
0 & 0\\
0 & \frac{1}{m}\\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0 \\
0 & 0 \\
0 & 0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0 & 0\\
0 & 0\\
0 & 0\\
0 & 0
\end{matrix}
 \right],
 \quad \textrm{rank}(\mathcal{C})=4,
\end{equation}
and thus all four states are controllable. This section starts by proving independent position control of many robots is not possible, but the mean position can be controlled. We then provide conditions under which the variance of many robots is also controllable.

\subsection{Independent control of many robots is impossible}
A single robot is fully controllable. For holonomic robots, movement in the $x$ and $y$ coordinates are independent, so for notational convenience without loss of generality we will focus only on movement in the $x$ axis. Given $n$ robots to be  controlled in the $x$ axis, there are $2n$ states: $n$ positions and $n$ velocities.
Our state-space representation is:
\begin{equation}
\begin{bmatrix}
\dot{x}_{1}\\ 
\ddot{x}_{1}\\
\vdots\\
\dot{x}_{n}\\
\ddot{x}_{n}

\end{bmatrix} = \begin{bmatrix}
0 & 1 & \ldots & 0 & 0 \\
0 & 0 & \ldots& 0 & 0 \\
\vdots &  \vdots & \ddots & \vdots & \vdots \\
0 & 0  & \ldots & 0 & 1 \\
0 & 0 & \ldots& 0 & 0 
\end{bmatrix}  \begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots \\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} + \begin{bmatrix}
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{equation}
 However, just as with one robot, we can only control two states because the controllability matrix $\mathcal{C}_n$ has rank two:
\begin{equation}
\mathcal{C}_n=\left[ \begin{matrix} 
0\\
1\\
\vdots\\
0\\
1
\end{matrix}
\,\middle\vert\,
  \begin{matrix} 
1\\
0\\
\vdots\\
1\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}
\,\middle\vert\,
\begin{matrix} 
0\\
0\\
\vdots\\
0\\
0
\end{matrix}\,\middle\vert\,
\ldots \right], \quad \textrm{rank}(\mathcal{C}_n)=2.
\end{equation}  
\subsection{Controlling the mean position of many robots}\label{sec:controlMeanPosition}
This means \emph{any} number of robots controlled by a global command have just two controllable states in each axis. We cannot arbitrarily control the position and velocity of two or more robots, but have options on which states to control.  One option is to control the position and velocity of the $j^{th}$ robot. To find a potentially more useful option, we create the following reduced order system that represents the average $x$ position and velocity of the $n$ robots:
\begin{align}
\begin{bmatrix}\nonumber
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} &= \frac{1}{n} \begin{bmatrix}
0& 1& \ldots &0& 1 \\
0& 0& \ldots &0& 0
\end{bmatrix}
\begin{bmatrix}
x_{1}\\
\dot{x}_{1}\\
\vdots\\
x_{n}\\
\dot{x}_{n}
\end{bmatrix} 
+ \frac{1}{n}\begin{bmatrix}
0& 0&  \ldots &0& 0 \\
0& 1&  \ldots &0& 1
\end{bmatrix}\begin{bmatrix} 
0\\
1\\
\vdots\\
0\\
1
\end{bmatrix} u_x
\end{align}
Thus:
\begin{equation}
\begin{bmatrix}
\dot{\bar{x}} \\
\ddot{\bar{x}}
\end{bmatrix} = \begin{bmatrix}
0& 1 \\
0& 0
\end{bmatrix}
\begin{bmatrix}
\bar{x}\\
\dot{\bar{x}}
\end{bmatrix} + \begin{bmatrix} 
0\\
1
\end{bmatrix} u_x
\end{equation}

We again analyze the controllability matrix $\mathcal{C}_{\mu}$:
\begin{equation}
\mathcal{C}_\mu=\left[ \begin{matrix} 
0\\
1
\end{matrix}
\,\middle\vert\,
 \begin{matrix} 
1\\
0
\end{matrix}
 \right],  \quad \textrm{rank}(\mathcal{C}_\mu)=2.
\end{equation}
Thus the average position and average velocity are controllable.

@@

We control mean position with a PD controller that uses the mean position and mean velocity. $[u_x,u_y]^\top$ is the global force applied to each robot:
\begin{align}
u_x &= K_{p}(x_{goal} - \bar{x}) + K_{d}(0-\dot{\bar{x}}) \nonumber\\
u_y &= K_{p}(y_{goal}  - \bar{y}) + K_{d}(0-\dot{\bar{y}})  \label{eq:PDcontrolPosition}
\end{align}
here $K_{p}$ is the proportional gain, and $K_{d}$ is the derivative gain.

@@

Controlling the variance requires being able to increase and decrease the variance.  We will list a sufficient condition for each. 
Microscale systems are affected by unmodelled dynamics. 
These unmodelled dynamics are dominated by Brownian noise, described by~\cite{einstein1956investigations}. 
To model this~\eqref{eq:stdform} must be modified as follows:
\begin{align}
\dot{\mathbf{x}}(t)  &=  A \mathbf{x}(t) + B \mathbf{u}(t) + W \bm{\varepsilon}(t),
\end{align}
where $W\bm{\varepsilon}(t)$ is a random perturbation produced by Brownian noise with magnitude $W$. Given a large obstacle-free workspace, a \emph{Brownian noise} process increases the variance linearly with time.
\begin{equation}
\dot{\sigma}_x^2(\mathbf{x}(t), \mathbf{u}(t) \Leftarrow 0)  = W \bm{\varepsilon},
\quad \sigma_x^2(t)  = \sigma_x^2(0) + W \bm{\varepsilon} t
\end{equation}
 If faster dispersion is needed, the swarm can be pushed through obstacles such as a diffraction grating or Pachinko board as in~\cite{Becker2013b}.

If robots with radius $r$ are in a bounded environment with sides of length $[\ell_x, \ell_y]$, the unforced variance asymptotically grows to the variance of a uniform distribution,
\begin{align}
[\sigma_x^2,\sigma_y^2] = \frac{1}{12}[ (\ell_x - 2 r)^2,(\ell_y - 2 r)^2].\label{eq:VarianceUniformDistribution}
\end{align}

@@

A  controller to regulate the variance to $\sigma^2_{ref}$ is:
\begin{align}
u_x &= K_{p}(x_{goal}(\sigma^2_{ref}) - \bar{x}) - K_{d}\bar{v}_x + K_{i}(\sigma^2_{ref}-\sigma^2_{x}) \nonumber\\
u_y &= K_{p}(y_{goal}(\sigma^2_{ref})  - \bar{y}) - K_{d}\bar{v}_y + K_{i}(\sigma^2_{ref}-\sigma^2_{y}).  \label{eq:PDcontrolVariance}
\end{align}
We call the gain scaling the variance error $K_i$ because the variance, if unregulated, integrates over time.
Eq.~\eqref{eq:PDcontrolVariance} assumes the nearest wall is to the left of the robot at $x=0$, and chooses a reference goal position that in steady-state would have the correct variance according to \eqref{eq:VarianceUniformDistribution}:
\begin{align}
x_{goal}(\sigma^2_{ref}) = \ell_x/2 = r + \sqrt{3\sigma^2_{ref}}
\end{align}
 If a wall to the right is closer, the signs of $[K_p,K_i]$ are inverted, and the location $x_{goal}$ is translated.

@@

\begin{algorithm}
\caption{Hybrid mean and variance control}\label{alg:MeanVarianceControl}
\begin{algorithmic}[1]
\Require Knowledge of swarm mean $[\bar{x},\bar{y}]$, variance $[\sigma_x^2, \sigma_y^2]$, the locations of the rectangular boundary $\{x_{min}, x_{max}, y_{min}, y_{max}\}$, and the target mean position $[x_{target},y_{target}]$.%TODO: use  \AND, \OR, \XOR, \NOT, \TO, \TRUE, \FALSE \gets
\State $flag_x \gets \FALSE$,  $flag_y \gets \FALSE$  \Comment{initially not in variance control mode}
\State $x_{goal} \gets  x_{target}$, $y_{goal} \gets y_{target}$
\Loop
%\State  Compute $\sigma_x^2, \sigma_y^2$
